% !TeX program = xelatex
\documentclass[12pt, oneside]{report}

% ==================== PACKAGES ====================
\usepackage[paper=a4paper, left=4cm, right=2cm, top=3cm, bottom=3cm]{geometry}
\usepackage{setspace}
\usepackage{fontspec}
\usepackage{titlesec}
\usepackage{fancyhdr}
\usepackage{tocloft}
\usepackage{blindtext} % remove after testing
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{microtype}

% Allow extra inter-word stretch as last resort to avoid overfull boxes
\emergencystretch=2em

% ==================== FONTS ====================
\setmainfont{Times New Roman}

% ==================== SPACING & PAGE STYLE ====================
\doublespacing
\pagestyle{fancy}
\fancyhf{}
\cfoot{\thepage}
\renewcommand{\headrulewidth}{0pt}

% ==================== CHAPTER & SECTION STYLES ====================
\titleformat{\chapter}{\normalfont\huge\bfseries}{\thechapter.}{1em}{}
\titleformat{\section}{\normalfont\large\bfseries}{\thesection}{1em}{}
\titleformat{\subsection}{\normalfont\normalsize\bfseries}{\thesubsection}{1em}{}

% ==================== TITLE PAGES & PRELIMINARIES ====================
\begin{document}

% Blank first leaf (thick paper)
\thispagestyle{empty}
\newpage

% ==================== OUTER COVER ====================
\input{frontmatter/title_outer.tex}

% ==================== INNER COVER ====================
\input{frontmatter/title_inner.tex}

% ==================== DECLARATION ====================
\input{frontmatter/declaration.tex}

% ==================== SUPERVISOR CERTIFICATION ====================
\input{frontmatter/certification.tex}

% Start Roman page numbering
\pagenumbering{roman}

% ==================== TABLE OF CONTENTS ====================
\tableofcontents
\newpage

% ==================== LISTS (if needed) ====================
\listoftables
\listoffigures

% ==================== ACKNOWLEDGEMENT ====================
\input{frontmatter/acknowledgement.tex}

% ==================== ABSTRACT ====================
\input{frontmatter/abstract.tex}

% Switch to Arabic numbering from Chapter 1
\pagenumbering{arabic}
\setcounter{page}{1}

% ==================== CHAPTERS ====================
\input{chapters/01_introduction.tex}
\input{chapters/02_literature_review.tex}
% ... add more chapters

% ==================== REFERENCES ====================
\clearpage
\addcontentsline{toc}{chapter}{References}
\begin{thebibliography}{99}

\bibitem{ahmad2011predicting}
Ahmad, H., Daud, A., Wang, L., Hong, H., Dawood, H. and Ahmad, W. (2011)
`Prediction of winning team in one day international cricket matches using machine learning'.
\textit{Journal of Computers}, 6(5), pp.~957--964.

\bibitem{bailey2004predicting}
Bailey, M. and Clarke, S.R. (2004)
`Predicting the match outcome in one day international cricket matches, while the match is in progress'.
\textit{Journal of Sports Science and Medicine}, 3(1), pp.~16--24.

\bibitem{brown2020language}
Brown, T., Mann, B., Ryder, N., Subbiah, M., Kaplan, J.D., Dhariwal, P., Neelakantan, A., Shyam, P., Sastry, G. and Askell, A. et al. (2020)
`Language models are few-shot learners'.
\textit{Advances in Neural Information Processing Systems}, 33, pp.~1877--1901.

\bibitem{bukiet2016evaluating}
Bukiet, B. and Harold, E.R. (2016)
`A Markov chain approach to baseball'.
\textit{Operations Research}, 45(1), pp.~14--23.

\bibitem{cricsheet}
Cricsheet (2024)
\textit{Cricsheet: A cricketing archive} [online].
Available at: \texttt{https://cricsheet.org} (Accessed: 18 April 2026).

\bibitem{davis2015markov}
Davis, J., Perera, H. and Swartz, T.B. (2015)
`Player evaluation in Twenty20 cricket'.
\textit{Journal of Sports Analytics}, 1(1), pp.~19--31.

\bibitem{dettmers2023qlora}
Dettmers, T., Pagnoni, A., Holtzman, A. and Zettlemoyer, L. (2023)
`QLoRA: Efficient finetuning of quantized LLMs'.
\textit{Advances in Neural Information Processing Systems}, 36, pp.~10088--10115.

\bibitem{devlin2019bert}
Devlin, J., Chang, M.W., Lee, K. and Toutanova, K. (2019)
`BERT: Pre-training of deep bidirectional transformers for language understanding'.
\textit{Proceedings of the 2019 Conference of the North American Chapter of the Association for Computational Linguistics}, pp.~4171--4186.

\bibitem{duckworth1998fair}
Duckworth, F.C. and Lewis, A.J. (1998)
`A fair method for resetting the target in interrupted one-day cricket matches'.
\textit{Journal of the Operational Research Society}, 49(3), pp.~220--227.

\bibitem{gruver2023large}
Gruver, N., Finzi, M., Qiu, S. and Wilson, A.G. (2023)
`Large language models are zero-shot time series forecasters'.
\textit{Advances in Neural Information Processing Systems}, 36.

\bibitem{haghighat2013review}
Haghighat, M., Rastegari, H. and Nourafza, N. (2013)
`A review of data mining techniques for result prediction in sports'.
\textit{ACSIJ Advances in Computer Science: an International Journal}, 2(5), pp.~7--12.

\bibitem{hegselmann2023tabllm}
Hegselmann, S., Buendia, A., Lang, H., Agrawal, M., Jiang, X. and Sontag, D. (2023)
`TabLLM: Few-shot classification of tabular data with large language models'.
\textit{Proceedings of the 26th International Conference on Artificial Intelligence and Statistics}, pp.~5549--5581.

\bibitem{hu2022lora}
Hu, E.J., Shen, Y., Wallis, P., Allen-Zhu, Z., Li, Y., Wang, S., Wang, L. and Chen, W. (2022)
`LoRA: Low-rank adaptation of large language models'.
\textit{International Conference on Learning Representations}.

\bibitem{huckerby2024llm}
Huckerby, J., Whitworth, A. and Dhaliwal, R. (2024)
`Leveraging large language models for football match outcome prediction'.
\textit{Machine Learning and Data Mining for Sports Analytics Workshop, ECML PKDD}.

\bibitem{iyer2009prediction}
Iyer, S.R. and Sharda, R. (2009)
`Prediction of athletes performance using neural networks: An application in cricket team selection'.
\textit{Expert Systems with Applications}, 36(3), pp.~5510--5522.

\bibitem{kampakis2015using}
Kampakis, S. and Thomas, W. (2015)
`Using machine learning to predict the outcome of English county twenty over cricket matches'.
\textit{Journal of Sports Analytics}, 1(1), pp.~63--76.

\bibitem{kirkpatrick2017overcoming}
Kirkpatrick, J., Pascanu, R., Rabinowitz, N., Veness, J., Desjardins, G., Rusu, A.A., Milan, K., Quan, J., Ramalho, T. and Grabska-Barwinska, A. et al. (2017)
`Overcoming catastrophic forgetting in neural networks'.
\textit{Proceedings of the National Academy of Sciences}, 114(13), pp.~3521--3526.

\bibitem{ouyang2022training}
Ouyang, L., Wu, J., Jiang, X., Almeida, D., Wainwright, C., Mishkin, P., Zhang, C., Agarwal, S., Slama, K. and Ray, A. et al. (2022)
`Training language models to follow instructions with human feedback'.
\textit{Advances in Neural Information Processing Systems}, 35, pp.~27730--27744.

\bibitem{perera2012markov}
Perera, H. and Swartz, T.B. (2012)
`Resource estimation in T20 cricket'.
\textit{IMA Journal of Management Mathematics}, 24(3), pp.~337--347.

\bibitem{preston2000batting}
Preston, I. and Thomas, J. (2000)
`Batting strategy in limited overs cricket'.
\textit{Journal of the Royal Statistical Society: Series D (The Statistician)}, 49(1), pp.~95--106.

\bibitem{rafailov2023direct}
Rafailov, R., Sharma, A., Mitchell, E., Manning, C.D., Ermon, S. and Finn, C. (2023)
`Direct preference optimization: Your language model is secretly a reward model'.
\textit{Advances in Neural Information Processing Systems}, 36.

\bibitem{rajapaksha2024transformer}
Rajapaksha, P., Aponso, A.C. and Bhatt, C. (2024)
`Transformer-based models for ball-by-ball outcome prediction in T20 cricket'.
\textit{International Journal of Sports Science \& Coaching}.

\bibitem{sankaranarayanan2014cricket}
Sankaranarayanan, V.V., Sattar, J. and Lakshminarayanan, V. (2014)
`Auto-play: A data mining approach to ODI cricket simulation and prediction'.
\textit{Proceedings of the SIAM International Conference on Data Mining}, pp.~1064--1072.

\bibitem{stern2016duckworth}
Stern, S.E. (2016)
`The Duckworth-Lewis-Stern method: Extending the Duckworth-Lewis methodology to deal with modern scoring rates'.
\textit{Journal of the Operational Research Society}, 67(12), pp.~1469--1480.

\bibitem{touvron2023llama}
Touvron, H., Lavril, T., Izacard, G., Martinet, X., Lachaux, M.A., Lacroix, T., Rozi\`{e}re, B., Goyal, N., Hambro, E. and Azhar, F. et al. (2023)
`LLaMA: Open and efficient foundation language models'.
\textit{arXiv preprint arXiv:2302.13971}.

\bibitem{vaswani2017attention}
Vaswani, A., Shazeer, N., Parmar, N., Uszkoreit, J., Jones, L., Gomez, A.N., Kaiser, {\L}. and Polosukhin, I. (2017)
`Attention is all you need'.
\textit{Advances in Neural Information Processing Systems}, 30.

\bibitem{vonwerra2022trl}
von Werra, L., Belkada, Y., Tunstall, L., Beeching, E., Thrush, T. and Lambert, N. (2022)
\textit{TRL: Transformer Reinforcement Learning} [online].
Available at: \texttt{https://github.com/huggingface/trl} (Accessed: 18 April 2026).

\bibitem{wei2022chain}
Wei, J., Wang, X., Schuurmans, D., Bosma, M., Chi, E., Le, Q. and Zhou, D. (2022)
`Chain-of-thought prompting elicits reasoning in large language models'.
\textit{Advances in Neural Information Processing Systems}, 35, pp.~24824--24837.

\end{thebibliography}

% ==================== APPENDICES ====================
\appendix
\input{appendices/appendix_a.tex}
% ... more appendices

% Blank last leaf
\clearpage
\thispagestyle{empty}
\newpage

\end{document}